Representing Scenes as Light Fields for Real Time Rendering in Virtual Reality Display

### 2021-03-09 TOG/rebuttal.md

# Grouped and Organized
## Shared
1. About hardware requirement (r1.1, r2.2):
   (r1) experiment is done on high-end machine but paper is targeting for mobile device?
   (r2) distance towards low latency, and low-storage VR/AR platforms?
2. About static scene (r1.2, r1.3, r3.5):
   (r1) no interaction, mobile devices can store
   (r3) scope should be clarified
3. About encoding performance (r3.5, r4.3):
   (r3) scope should be clarified
   (r4) viewing is instant, but creation is not (encoding/training is slow)
4. About user study (r1.4, r3.4, r4.6):
   (r1, r4) short, gaze fixed / static
   (r3) narrow foveation methods selection
5. About constraint (r1.6, r2.5, r2.6, r4.4, r4.7):
   (r1) restricted position and direction?
   (r2) low (angular) resolution // head volume? frequency of transmiting network?
   (r4) occlusion (motion parallel): not clearly demonstrated
6. Publish source code (r3.3, r5.1):
   (r5) to help understanding
7. About written (r3.1, r3.2, r4.14):
   (r3) improve clarity / bib consistency
   (r4) typo / style

## Individual
Review 1
* Foveation rendering as ground truth (r1.5)

Review 2
* Concern about the comparison with MSI ([Broxton2020] or [Attal2020]):
  MSI can achieve same low storage under constraint of no view-dependent effects, small motion range, low resolution (r2.4, r2.6, r2.7)
* Angular resolution: 6.4 PPD vs ground truth 13 PPD? (r2.1)
* Experiment about storage? (r2.3)

Review 3 (None)

Review 4
* not comparable between nerual and mesh representation (r4.2)
* A fairer comparison: allow NeRF an additional 96 channels (r4.8)
* Why our is better? foveat network: still represents high-res whole scene, why baseline is worse? (4.9)
* Artifacts in video:
  1m30: double image artifacts (r4.5)
  1m03: black splotches, where they come from? (r4.10)
* Minor questions:
  Misunderstanding: need depth reconstruction for 360 photos application. (r4.1)
  Not understande Fig 7 (r4.11)
  Isn't the representation just alpha composited front to back? (r4.12)
  Is NeRF here also running on a TensorRT implementation? (r4.13)

Review 5
* Hard to read (Not understand)
  "The entire experiment of each user consists of 36 trials, 6 trials per pair of ordered conditions."
  --> how did this come about? 3 pairs of conditions x 2 scenes? (r5.2)
  "Each condition from an individual scene consists of 3 different gaze positions."
  --> I did not understand what this meant. Once participants had moved their eyes 3 times, the experiment stopped? (r5.3)
  "LPIPS uses deep neural networks to estimate perceptual similarities between the image provided and a reference image."
  --> What is meant by perceptual similarity? color similarity? (r5.4)
  "We used one-way repeated measures ANOVAs to compare effects across three stimuli on each eccentricity range and together."
  --> What is being treated as repeated measures? what is the independent and dependent variable? and what does "together" refer to? (r5.5)

# Raw Collections
## Review 1
1. Target mobile device but experiment on high-end machine?
2. Static scene: mobile devices can store. how about dynamic scene?
3. No interaction
4. user study: fix gaze, short
5. foveation rendering as ground truth
6. restricted position and direction?

## Review 2
1. Angular resolution: 6.4 PPD vs ground truth 13 PPD?
2. Performance: hardware? distance from low latency, and low-storage VR/AR platforms?
3. experiment about storage?
4. MSI: 10PPD - 10 MB -> 6PPD - 0.06MB?
5. head volume? frequency of transmiting network?
6. powerful computer but low resolution. MSI can achieve same storage real-time on mobile
7. why ray-caster? why not bake to MSI? (Under constraint of no view-dependent effects, small motion range, low resolution)

## Review 3
1. Improvement of clarity: the introduction of the egocentric coordinates might be a bit dense/slow to read
2. Reference: make bibliography more consistent
3. Publish source code
4. User study: the selection of foveation methods to compare is somewhat narrow
5. The scope should be clearly stated: static and receive side (decoder) only

## Review 4
1. Misunderstanding: 360 photos application: need depth reconstruction, e.g., Serrano et al. 2019 IEEE VR.
2. not comparable between nerual and mesh representation
3. Encode is slow, viewing is instant, but creation is not
4. Occlusion: small.
5. double image artifacts at 1m30s in video
6. Static user study: only a subset of experience
7. Translation and occlusion (motion parallax) are not clearly demonstrated and validated
8. A fairer comparison: allow NeRF an additional 96 channels
9. Why our is better? foveat network: still represents high-res whole scene, why baseline is worse?
10. Artifacts: 1m03 in video: black splotches, where they come from?
11. (Minor) Not understande Fig 7
12. (Minor) Isn't the representation just alpha composited front to back?
13. (Minor) Is NeRF here also running on a TensorRT implementation?
14. Typos/style:
    L154: "remarkably" <- can remove this; the reader can determine whether it is remarkable or not.
    L324: using 'd' for density here is not a huge problem but might be confused for 'depth'. Later on (L614) this is refered ot as 'a'.
    L1064: "unprecedented performance" <- please tone it down a bit; every technique is a trade-off somewhere.
    Duplicated citation of Attal et al.

## Review 5
1. Hard to read
2. "The entire experiment of each user consists of 36 trials, 6 trials per pair of ordered conditions." --> how did this come about? 3 pairs of conditions x 2 scenes?
3. "Each condition from an individual scene consists of 3 different gaze positions." --> I did not understand what this meant. Once participants had moved their eyes 3 times, the experiment stopped?
4. "LPIPS uses deep neural networks to estimate perceptual similarities between the image provided and a reference image." What is meant by perceptual similarity? color similarity?
5. "We used one-way repeated measures ANOVAs to compare effects across three stimuli on each eccentricity range and together." --> Unfortunately this is too concise an explanation to make sense -- what is being treated as repeated measures? what is the independent and dependent variable? and what does "together" refer to?

### 2021-03-09 TOG/0-abstract.tex
submission

Tailored for the future portable, low-storage, and energy-efficient VR platforms, we present the first gaze-contingent 3D neural representation and view synthesis method. We incorporate the human psychophysics of visual- and stereo-acuity into egocentric neural representation of 3D scenery. Furthermore, we jointly optimize the latency/performance and visual quality, while mutually bridging human perception and neural scene synthesis, to achieve perceptually high quality immersive interaction. Both objective analysis and subjective study demonstrate the effectiveness of our approach in significantly reducing local storage volume and synthesis latency (up to 99\% reduction in both data size and computational time), while simultaneously presenting high-fidelity rendering, with perceptual quality identical to that of fully locally stored and rendered high-quality imagery.
\end{abstract}

>>>>

Tailored for the future portable, low-storage, and energy-efficient VR platforms, we present the first gaze-contingent 3D neural representation and view synthesis method. We incorporate the human psychophysics of visual- and stereo-acuity into an egocentric neural representation of 3D scenery. Furthermore, we jointly optimize the latency/performance and visual quality, while mutually bridging human perception and neural scene synthesis, to achieve perceptually high-quality immersive interaction. Both objective analysis and subjective study demonstrate the effectiveness of our approach in significantly reducing local storage volume and synthesis latency (up to 99\% reduction in both data size and computational time), while simultaneously presenting high-fidelity rendering, with perceptual quality identical to that of fully locally stored and rendered high-quality imagery.
\end{abstract}

### 2021-03-09 TOG/1-introduction.tex
submission

\note{
What problem we are trying to solve.
Why it is important, and why people should care.
}%note
Efficient storage is a core challenge faced by today's computational systems, especially due to the fast-growing high resolution acquisition/display devices, the demand for 3D data quality, and machine learning.
Thus, streaming services and remote cloud-based rendering (such as with Google Stadia and Nvidia Shield) have rapidly emerged in recent years. These storage-friendly systems have shown an unprecedented potential to reshape the traditional computer graphics pipeline by completely relieving the end user's computational workload and storage pressure, thus advancing device portability. Future spatial and immersive computing platforms such as VR and AR are potentially major beneficiaries due to the extensive demands of portability and energy-efficiency  \footnote{\url{https://www.vxchnge.com/blog/virtual-reality-data-storage-concern}}.

However, existing low-storage services have only been successfully deployed in relatively latency-insensitive platforms such as TV-based gaming consoles, due to the inevitable latency from transmission (more than 100ms \footnote{\url{https://www.pcgamer.com/heres-how-stadias-input-lag-compares-to-native-pc-gaming/}}). 
In comparison, VR and AR require remarkably low latency and significantly higher fidelity/speed for sickness-free and natural experience. As a consequence, there is a demand for massive high-quality spatial-temporal data. Unfortunately, no solution exists today to deliver high-quality and dynamic/interactive immersive content at low latency and with low local storage.

@@

Recent years have witnessed a surge in neural representations as an alternative to compact 3D representations such as point clouds and polygonal meshes, enabling compact assets. Application of neural representations include voxelization \cite{sitzmann2019deepvoxels}, distance fields \cite{sitzmann2019metasdf,mildenhall2020nerf,park2019deepsdf} and many more. However, existing solutions suffer from either high time consumption or low image fidelity with large scenes.

Acceleration without compromising perceptual quality is also an ultimate goal of real-time rendering.
Gaze-contingent rendering approaches have shown remarkable effectiveness of presenting imagery of perceptual quality identical to full resolution rendering~\cite{Guenter:2012:F3G,Patney:2016:TFR,Tursun:2019:LCA}. This is achieved by harnessing high field-of-view(FoV) head-mounted displays and high precision eye-tracking technologies. However, existing foveated rendering approaches still require locally stored full 3D assets, resulting in long transmission time and storage volume, whose availability is unfortunately limited in wearable devices.

@@

To overcome these open challenges mentioned above, we present the first gaze-contingent and interactive neural synthesis for high-quality 3D scenes. Aiming at the next-generation wearable, low-latency and low-storage VR/AR platforms, our compact neural model predicts stereo frames given users' real-time head and gaze motions. The predicted frames are perceptually identical to traditional real-time rendering of high-quality scenes.
With wide FoV and high resolution eye-tracked VR displays, our method significantly reduces 3D data storage (from 100MB to 0.5MB) thus enables instant viewing (without long loading/transmission waits) compared to traditional rendering pipelines; preserves visual quality, and improves responsiveness (from 9s to as low as 20 ms) compared with alternative view neural synthesis methods. 
Tailoring to future lightweight, power and memory-efficient wearable display systems, expected to be active all-day long akin to today's everyday-use mobile phones, our method achieves minimal local storage for high-quality 3D scenes.  %\praneeth{---- somoene please check the text in the brackets and make sure it is ok :)}

@@

To achieve immersive viewing with low storage, high visual quality as well as low systematic latency, we introduce spatial-angular human psychophysical characteristics to neural representation and synthesis. 
%To this end, we build both the retina-matched neural models and VR-tailored representation coordinates, towards high visual quality as well as low systematic latency. \praneeth{-- this line is a little confusing to me}. 
Specifically, we represent 3D scenes with concentric spherical coordinates. Our representation also allows for depicting the foveated color sensitivity and stereopsis, during both the training and inference phases. 
Furthermore, we derive an analytical spatial-temporal perception model to optimize our neural scene representation towards imperceptible loss in image quality and latency.\qisun{better? This paragraph was a bit messy before. Tried to clean up.}

@@

We validate our approach by conducting numerical analysis and user evaluations on commercially available AR/VR display devices. The series of experiments reveal our method's effectiveness in delivering an interactive VR viewing experience, perceptually identical to an arbitrarily high quality rendered 3D scene, with significantly lowered data storage. We make the following major contributions:
\begin{itemize}
    \item A low-latency, interactive, and significantly low-storage immersive application. It also offers full perceptual quality with instant high resolution and high FoV first-person VR viewing.
    \item A scene parameterization and vision-matched neural representation/synthesis method considering both visual- and stereo- acuity.
    \item A spatiotemporally analytical model for jointly optimizing systematic latency and perceptual quality.
\end{itemize}

>>>>

\note{
What problem we are trying to solve.
Why it is important, and why people should care.
}%note
Efficient storage is a core challenge faced by today's computational systems, especially due to the fast-growing high-resolution acquisition/display devices, the demand for 3D data quality, and machine learning.
Thus, streaming services and remote cloud-based rendering (such as Google Stadia and Nvidia Shield) have rapidly emerged in recent years. These storage-friendly systems have shown an unprecedented potential to reshape the traditional computer graphics pipeline by completely relieving the end user's computational workload and storage pressure, thus advancing device portability. Future spatial and immersive computing platforms such as VR and AR are potentially major beneficiaries due to the extensive demands of portability and energy-efficiency  \footnote{\url{https://www.vxchnge.com/blog/virtual-reality-data-storage-concern}}.

However, existing low-storage services have only been successfully deployed in relatively latency-insensitive platforms such as TV-based gaming consoles due to the inevitable latency from transmission (more than 100ms \footnote{\url{https://www.pcgamer.com/heres-how-stadias-input-lag-compares-to-native-pc-gaming/}}). 
In comparison, VR and AR require remarkably low latency and significantly higher fidelity/speed for a sickness-free and natural experience. As a consequence, there is a demand for massive high-quality spatial-temporal data. Unfortunately, no solution exists today to deliver high-quality and dynamic/interactive immersive content at low latency and with low local storage.

@@

Recent years have witnessed a surge in neural representations as an alternative to compact 3D representations such as point clouds and polygonal meshes, enabling compact assets. Application of neural representations includes voxelization \cite{sitzmann2019deepvoxels}, distance fields \cite{sitzmann2019metasdf,mildenhall2020nerf,park2019deepsdf}, and many more. However, existing solutions suffer from either high time consumption or low image fidelity with large scenes.

Acceleration without compromising perceptual quality is also an ultimate goal of real-time rendering.
Gaze-contingent rendering approaches have shown remarkable effectiveness in presenting imagery of perceptual quality identical to full resolution rendering~\cite{Guenter:2012:F3G,Patney:2016:TFR,Tursun:2019:LCA}. This is achieved by harnessing high field-of-view(FoV) head-mounted displays and high precision eye-tracking technologies. However, existing foveated rendering approaches still require locally stored full 3D assets, resulting in long transmission time and storage volume, whose availability is unfortunately limited in wearable devices.

@@

To overcome these open challenges mentioned above, we present the first gaze-contingent and interactive neural synthesis for high-quality 3D scenes. Aiming at the next-generation wearable, low-latency, and low-storage VR/AR platforms, our compact neural model predicts stereo frames given users' real-time head and gaze motions. The predicted frames are perceptually identical to the traditional real-time rendering of high-quality scenes.
With wide FoV and high-resolution eye-tracked VR displays, our method significantly reduces 3D data storage (from 100MB mesh to 0.5MB neural model) compared with traditional rendering pipelines; it also remarkably improves responsiveness (from 9s to 20 ms) compared with alternative view neural synthesis methods. That is, we enable instant immersive viewing without long loading/transmission waits while preserving visual fidelity.
Tailored to future lightweight, power and memory-efficient wearable display systems, expected to be active all-day long akin to today's everyday-use mobile phones, our method achieves minimal local storage for high-quality 3D scenes.  %\praneeth{---- somoene please check the text in the brackets and make sure it is ok :)}

@@

To achieve immersive viewing with low storage, high visual quality and low systematic latency, we introduce spatial-angular human psychophysical characteristics to neural representation and synthesis. 
%To this end, we build both the retina-matched neural models and VR-tailored representation coordinates, towards high visual quality as well as low systematic latency. \praneeth{-- this line is a little confusing to me}. 
Specifically, we represent 3D scenes with concentric spherical coordinates. Our representation also allows for depicting the foveated color sensitivity and stereopsis during both the training and inference phases. 
Furthermore, we derive an analytical spatial-temporal perception model to optimize our neural scene representation towards imperceptible loss in image quality and latency.\qisun{better? This paragraph was a bit messy before. Tried to clean up.}

@@

We validate our approach by conducting numerical analysis and user evaluation on commercially available AR/VR display devices. The series of experiments reveal our method's effectiveness in delivering an interactive VR viewing experience, perceptually identical to an arbitrarily high quality rendered 3D scene, with significantly lowered data storage. We will open source our implementation and dataset.
We make the following major contributions:
\begin{itemize}
    \item A low-latency, interactive, and significantly low-storage immersive application. It also offers full perceptual quality with instant high resolution and high FoV first-person VR viewing.
    \item A scene parameterization and vision-matched neural representation/synthesis method considering both visual- and stereo- acuity.
    \item A spatiotemporally analytical model for jointly optimizing systematic latency and perceptual quality.
\end{itemize}

### 2021-03-09 TOG/2-prior.tex
submission

\subsection{Implicit Scene Representation for Neural Rendering}
To fully represent a 3D object and environment without large explicit storage, neural representations have drawn extensive attentions in both computer graphics and computer vision communities. 
With a deep neural network that depicts a 3D world as an implicit function, the neural networks may directly approximate an object's appearance given a camera pose \cite{sitzmann2019deepvoxels,sitzmann2019srns,sitzmann2019siren,mildenhall2020nerf}. 
Prior arts also investigated on implicitly representing shapes ~\cite{park2019deepsdf} and surfaces ~\cite{mescheder2019occupancy} to improve the visual quality of 3D objects. Inspired by \cite{park2019deepsdf}, signed distance function has been deployed for representing raw data such as point cloud ~\cite{atzmon2020sal,gropp2020implicit} with better efficiency ~\cite{chabra2020deep}. 
The input information was varied from 2D images ~\cite{lin2020sdf, yariv2020multiview}, 3D context ~\cite{saito2019pifu, oechsle2019texture}, time-varying vector field~\cite{niemeyer2019occupancy}, and local features~\cite{tretschk2020patchnets, liu2020neural}.
Current implicit representations primarily focus on locally ``out-side-in'' viewing of individual objects, with low field-of-view, speed and resolution. For large scenes, the lack of coverage may cause quality drop.
However, in virtual reality, the cameras are typically first-person and highly dynamic, requiring low latency, high resolution, and 6DoF coverage. 
%Our method, in comparison, is VR tailored via representing the scene with user-centered views.

\subsection{Gaze-Contingent Rendering}
Both rendering and neural synthesis suffer from the heavy computational load, thus systematic latency and lags.
With a full mesh representation, foveated rendering has been proposed to accelerate local rendering performance, including mesh- \cite{Guenter:2012:F3G,Patney:2016:TFR} or image-based \cite{Sun:2017:PGF,Kaplanyan:2019:DNR} methods. The accelerations are typically achieved through high field-of-view displays (such as VR head-mounted displays) and eye-tracking technologies. However, these all require full access to the original 3D assets or per-frame images. Consequently, the systems usually require large local storage, long transmission waits, or constant frame-based transmissions with latency. 
%On the other hand, most view synthesizing methods are designed for offline viewing or processing, other than real-time or interactive consumption, especially with high resolution and FoV.  
Our method bridges human vision to deep synthesis. It accelerates neural synthesis beyond local rendering, enabling both computational and storage efficiency. It is suitable for future cloud-based virtual reality with instant full-quality viewing and interaction.
%\zh{seems not sufficient?}

\subsection{Panorama-Based 6DoF Immersive Viewing}
Panorama images and videos greatly fit VR applications thanks to their 360 FoV coverage. However, the main challenge is their single projection center in each frame, limiting free-form camera translation, thus 6DoF immersive viewing experience. 
Recently, extensive research has been proposed to address this problem. Serrano et al. \shortcite{Serrano_TVCG_VR-6dof} presented a depth-based dis-occlusion method that dynamically reprojects to 6DoF cameras, enabling natural viewing with motion parallax. Pozo et al. \shortcite{Pozo:2019:I6V} jointly optimizes capture and viewing processes.
Machine learning approaches have advanced the robustness of various geometric and lighting conditions \cite{Attal:2020:ECCV,Lin:DeepPanorama,Benjamin:2020:RTV}. 
As a spherical image, panoramas are designed for capturing physical worlds with flexible reprojections to displays. The 6DoF viewing typically suffer from trade-offs among performance, achievable resolution, and dis-occlusion effects due to lack of knowledge of the 3D space.
With an orthogonal mission, our model represents and reproduces arbitrary 3D virtual worlds for low storage VR. It lowers memory usage by replacing traditional mesh- or volume-based data than predicting image-space occlusions.

>>>>

\subsection{Implicit Scene Representation for Neural Rendering}
To fully represent a 3D object and environment without large explicit storage, neural representations have drawn extensive attention in computer graphics and computer vision communities. 
With a deep neural network that depicts a 3D world as an implicit function, the neural networks may directly approximate an object's appearance given a camera pose \cite{sitzmann2019deepvoxels,sitzmann2019srns,sitzmann2019siren,mildenhall2020nerf,park2017transformation}. 
Prior arts also investigated implicitly representing shapes ~\cite{park2019deepsdf} and surfaces ~\cite{mescheder2019occupancy} to improve the visual quality of 3D objects. Inspired by \cite{park2019deepsdf}, signed distance functions have been deployed for representing raw data such as point cloud ~\cite{atzmon2020sal,gropp2020implicit} with better efficiency ~\cite{chabra2020deep}. 
The input information ranged from 2D images ~\cite{lin2020sdf, yariv2020multiview,choi2019extreme}, 3D context ~\cite{saito2019pifu, oechsle2019texture, zhang2020deep}, time-varying vector field~\cite{niemeyer2019occupancy} to local features~\cite{tretschk2020patchnets, liu2020neural}.
Current implicit representations primarily focus on locally ``outside-in'' viewing of individual objects, with low field-of-view, speed, and resolution. For large scenes, the lack of coverage may cause the quality to drop.
However, in virtual reality, the cameras are typically first-person and highly dynamic, requiring low latency, high resolution, and 6DoF coverage. 
%Our method, in comparison, is VR tailored via representing the scene with user-centered views.

\subsection{Gaze-Contingent Rendering}
Both rendering and neural synthesis suffer from the heavy computational load, thus systematic latency and lags.
With a full mesh representation, foveated rendering has been proposed to accelerate local rendering performance, including mesh- \cite{Guenter:2012:F3G,Patney:2016:TFR} or image-based \cite{Sun:2017:PGF,Kaplanyan:2019:DNR} methods. The accelerations are typically achieved through high field-of-view displays (such as VR head-mounted displays) and eye-tracking technologies. However, these all require full access to the original 3D assets or per-frame images. Consequently, the systems usually require large local storage, long transmission waits, or constant frame-based transmissions with latency. 
%On the other hand, most view synthesizing methods are designed for offline viewing or processing, other than real-time or interactive consumption, especially with high resolution and FoV.  
Our method bridges human vision and deep synthesis. It accelerates neural synthesis beyond local rendering, enabling both computational and storage efficiency. It is also suitable for future cloud-based virtual reality with instant full-quality viewing and interaction.
%\zh{seems not sufficient?}

\subsection{Panorama-Based 6DoF Immersive Viewing}
Panorama images and videos much fit VR applications thanks to their 360 FoV coverage. However, the main challenge is their single projection center in each frame, limiting free-form camera translation, thus 6DoF immersive viewing experience. 
Recently, extensive research has been proposed to address this problem. Serrano et al. \shortcite{Serrano_TVCG_VR-6dof} presented a depth-based dis-occlusion method that dynamically reprojects to 6DoF cameras, enabling natural viewing with motion parallax. Pozo et al. \shortcite{Pozo:2019:I6V} jointly optimizes capture and viewing processes.
Machine learning approaches have advanced the robustness of various geometric and lighting conditions \cite{Attal:2020:ECCV,Lin:DeepPanorama,Benjamin:2020:RTV}. 
As a spherical image, panoramas are designed for capturing physical worlds with flexible reprojections to displays. The 6DoF viewing typically suffers from trade-offs among performance, achievable resolution, and dis-occlusion effects due to lack of knowledge of the 3D space.
With an orthogonal mission, our model represents and reproduces arbitrary 3D virtual worlds for low storage VR. It lowers memory usage by replacing traditional mesh- or volume-based data than predicting image-space occlusions.
